---
title: 'DATA 605'
author: "Vinayak Patel"
date: "Sept 8, 2019"
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
subtitle: 'Week 2: Assignment'
---

# Problem Set 1
1)
Show that $A^TA \neq AA^T$ in general. (Proof and demonstration.)

```{r}
A = matrix(seq(1, 9), nrow=3, byrow = T)
A
AT = matrix(seq(1, 9), nrow=3, byrow = F)
AT
```

```{r}
A %*% AT
```

```{r}
AT %*% A
```

```{r}
(AT %*% A) == (A %*% AT)
```

__Hence, $A^TA \neq AA^T$__

2)
For a special type of square matrix A, we get $A^T A = AA^T$ . Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).


$A^T A = A A^T$ is true if $A$ is a diagonal matrix.So An identity matrix when transposed and multiplied to itself, are equal to each other.

Let $A = \begin{bmatrix} 1 & 0 & 0 & 0\\0 & 1 & 0 &0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix}$ then
$A^T =  \begin{bmatrix} 1 & 0 & 0 & 0\\0 & 1 & 0 &0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix}$ therefore:

$$
A^T A =
\begin{bmatrix} 1 & 0 & 0 & 0\\0 & 1 & 0 &0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix} 
\begin{bmatrix} 1 & 0 & 0 & 0\\0 & 1 & 0 &0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix}  =
\begin{bmatrix} 1 & 0 & 0 & 0\\0 & 1 & 0 &0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix}
$$
$$
A A^T =
\begin{bmatrix} 1 & 0 & 0 & 0\\0 & 1 & 0 &0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix} 
\begin{bmatrix} 1 & 0 & 0 & 0\\0 & 1 & 0 &0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix}  =
\begin{bmatrix} 1 & 0 & 0 & 0\\0 & 1 & 0 &0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix}
$$
This gives a case when $A^T A = A A^T$ is true.

```{r}
A = matrix(c(1,1,0,-1,1,3,2,4,0,2,4,8,-1,4,8,6), nrow=2, byrow = T)
AT = t(A)
(A %*% AT) == (A %*% AT)
```


$A^TA=AA^T$, we can see that all of them are symmetric matrices. A symmetric matrix is a matrix where $A_{ij}=A_{ji}$ for every i and j.

---

#Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track ights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your ight using radars.
Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. Please submit your response in an R Markdown document using our class naming convention, E.g. LFulton_Assignment2_PS2.png You don't have to worry about permuting rows of A and you can assume that A is less
than 5x5, if you need to hard-code any variables in your code. If you doing the entire assignment in R, then please submit only one markdown document for both the problems.


Write a function to factorize a square matrix A into LU or LDU.
### Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer:
```{r}

#matrix factorization


LU_factorization <- function(A) {
  # Check wheter matrix is square or not
  if (dim(A)[1]!=dim(A)[2]) {
    return(NA)
  }
  
  U <- A
  n <- dim(A)[1]
  L <- diag(n)
  
  if (n==1) {
    return(list(L,U))
  }
  
  for(i in 2:n) {
    for(j in 1:(i-1)) {
      multiplier <- -U[i,j] / U[j,j]
      U[i, ] <- multiplier * U[j, ] + U[i, ]
      L[i,j] <- -multiplier
    }
  }
  return(list(L,U))
}
```


```{r}
# test matrix factorization
### Sample 1: A 2x2
A<- matrix(c(2,6,1,8), nrow = 2, ncol = 2)
LU <-LU_factorization(A)
L<-LU[[1]]  
U<-LU[[2]]
A
L
U
A==L%*%U

### Sample 2: B 3x3
B <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), nrow = 3, ncol = 3)
LU <- LU_factorization(B)
L<-LU[[1]]  
U<-LU[[2]]
B
L
U
B==L%*%U


### Sample 3: D 3x3 with lots of negative inputs

C <- matrix(c(-3,5,-1,4,-2,-6,7,1,8), nrow = 3, ncol = 3)
LU <-LU_factorization(C)
L<-LU[[1]]  
U<-LU[[2]]
C
L
U
C==L%*%U
```
__Hence, it verifies the result__


